Supplementary Information 1

### 2025-04-14 MS2_Krill.Rmd
adding conceptual mapa to MJ

Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under four modeling scenarios from 1990 to 2020. The reference model (blue), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015. In contrast, the models that incorporate ecosystem components—predator data (green), environmental data (yellow), or both (red)—estimate consistently lower spawning output throughout the time series. The red line, representing the model with both predator and environmental data, remains the most conservative, suggesting that ecosystem-informed models lead to more precautionary estimates of reproductive potential. The rigth panel displays the posterior density distributions of unfished spawning biomass (SSB~o~) for the same scenarios. The reference model estimates the highest SSBo values (mode ~ 3.2), while the ecosystem-informed models estimate considerably lower SSBo, with the most constrained distribution occurring under the full ecosystem model (red). The shift in the distributions indicates that including ecosystem drivers reduces estimates of unfished reproductive capacity, which has implications for reference point calculations and harvest strategies.

```{r compara,eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
knitr::include_graphics('Figs/COM1compare2_spawnbio_uncertainty.png')

@@

```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated spawning output for Antarctic krill under four assessment scenarios"}
knitr::include_graphics('Figs/COM1compare16_densities_SR_LN(R0).png')

>>>>

Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under most significant models in this analisys, s1.1 and s1.4. The reference model (red), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015. In contrast, the models that incorporate ecosystem components—predator data (green), environmental data (yellow), or both (red)—estimate consistently lower spawning output throughout the time series. The red line, representing the model with both predator and environmental data, remains the most conservative, suggesting that ecosystem-informed models lead to more precautionary estimates of reproductive potential. The rigth panel displays the posterior density distributions of unfished spawning biomass (SSB~o~) for the same scenarios. The reference model estimates the highest SSBo values (mode ~ 3.2), while the ecosystem-informed models estimate considerably lower SSBo, with the most constrained distribution occurring under the full ecosystem model (red). The shift in the distributions indicates that including ecosystem drivers reduces estimates of unfished reproductive capacity, which has implications for reference point calculations and harvest strategies.

```{r compara,eval=TRUE, out.width="60%", fig.cap=""}
knitr::include_graphics('Figs/alloutput_s1_4.png')

@@

```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
knitr::include_graphics('Figs/R0plot.png')

### 2025-04-14 index.Rmd
adding conceptual mapa to MJ

```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets. Differences among strata have been identified from various perspectives of krill population dynamics, as well as from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).

@@

## Statistical Model (SS3)

@@

```
Source of data input

```{r fig.width=9, fig.height=4}
dat1_std <- dat1$CPUE %>%
  mutate(
    index_std = (obs - min(obs)) / (max(obs) - min(obs)),
    index = recode(index,
                   `1` = "FISHERYBS",
                   `2` = "FISHERYEI",
                   `3` = "FISHERYGS",
                   `4` = "FISHERYJOIN",
                   `5` = "FISHERYSSIW",
                   `6` = "SURVEYBS",
                   `7` = "SURVEYEI",
                   `8` = "SURVEYGS",
                   `9` = "SURVEYJOIN",
                   `10` = "SURVEYSSIW",
                   `11` = "PREDATOR"),
    index = factor(index, levels = c("FISHERYBS", 
                                     "FISHERYEI", 
                                     "FISHERYGS",
                                     "FISHERYJOIN", 
                                     "FISHERYSSIW", 
                                     "SURVEYBS", 
                                     "SURVEYEI", 
                                     "SURVEYGS", 
                                     "SURVEYJOIN", 
                                     "SURVEYSSIW", 
                                     "PREDATOR")),
    category = case_when(
      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
      index == "PREDATOR" ~ "Predator",
      TRUE ~ NA_character_  
    )
  )

@@

```{r fig.width=10, fig.height=4, warning=FALSE}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

dat_expanded <- dat_long %>%
  uncount(weights = Frecuencia)

p2 <- ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
  geom_density_ridges_gradient(scale = 1.8,
                               rel_min_height = 0.001,
                               col = "black",
                               alpha=0.4) +
  facet_wrap(~FltSvy, ncol = 10) +
  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
             color = "red", 
             size = 0.7,
             alpha = 0.5) +
  scale_fill_manual(values = c("FISHERYBS" = '#1b9e77',  
                               "FISHERYEI" = '#1b9e77',  
                               "FISHERYGS" = '#1b9e77',
                               "FISHERYJOIN" = '#1b9e77',  
                               "FISHERYSSIW" = '#1b9e77',  
                               "SURVEYBS" = '#d95f02',  
                               "SURVEYEI" = '#d95f02',  
                               "SURVEYGS" = '#d95f02',  
                               "SURVEYJOIN" = '#d95f02',  
                               "SURVEYSSIW" = '#d95f02',  
                               "PREDATOR" = '#7570b3')) +  
  labs(title = "",
       x = "cm",
       y = "",
       fill = "Category") +
  theme_bw()+
  theme(legend.position = "none")
p2
```

@@

```{r}
# Crear una columna de modelo en cada data frame
outps1$Model <- "Ref Model: No Env-Predator"
outps2$Model <- "s1.2 w/ Predator data"
outps3$Model <- "s1.3 w/ Env data"
outps4$Model <- "s1.4 w/ Env and Predator data"

@@

outpsall_long <- outpsall %>%
  pivot_longer(cols = c(Bio_smry, SpawnBio, Recruit_0), 
               names_to = "Variable", values_to = "Value")

# Gráfico de barras agrupadas por variable y modelo
ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, 
             ncol=1,
             scales = "free_y") + 
  scale_fill_manual(name= "Scenario",
                    values = c("Ref Model: No Env-Predator" = "#ca0020",

                               "s1.4 w/ Env and Predator data" = "#404040")) +
  labs(title = "",
       x = "", y = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "bottom")
```

@@

SSplotComparisons(summaryoutput1_4,
                  legendlabels = c("Ref Model: No Env-Predator",
                 "S1.1 w/ Predator data",
                 "S1.1 w/ Env data",
                 "S1.1 w/ Env and Predator data"),
                 filenameprefix = "COM1",
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,
                 plotdir=Figs)
```

@@

>>>>

```{r mapa, out.width='50%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
knitr::include_graphics('Figs/481.png')
```

In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014]. This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).

@@

## Statistical Model (`SS3`)

@@

```
Source of data input.

Ibdex abundance in Figure \@ref(fig:index)

```{r}
dat1_std <- dat1$CPUE %>%
  mutate(
    index_std = (obs - min(obs)) / (max(obs) - min(obs)),
    index = recode(index,
                   `1` = "FISHERYBS",
                   `2` = "FISHERYEI",
                   `3` = "FISHERYGS",
                   `4` = "FISHERYJOIN",
                   `5` = "FISHERYSSIW",
                   `6` = "SURVEYBS",
                   `7` = "SURVEYEI",
                   `8` = "SURVEYGS",
                   `9` = "SURVEYJOIN",
                   `10` = "SURVEYSSIW",
                   `11` = "PREDATOR"),
    index = factor(index, levels = c("FISHERYBS", 
                                     "FISHERYEI", 
                                     "FISHERYGS",
                                     "FISHERYJOIN", 
                                     "FISHERYSSIW", 
                                     "SURVEYBS", 
                                     "SURVEYEI", 
                                     "SURVEYGS", 
                                     "SURVEYJOIN", 
                                     "SURVEYSSIW", 
                                     "PREDATOR")),
    category = case_when(
      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
      index == "PREDATOR" ~ "Predator",
      TRUE ~ NA_character_  
    )
  )
```

```{r index, fig.width=9, fig.height=4, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1. Each panel represents a distinct spatial or functional stratum, with trend lines indicating temporal variation from 1990 to 2020. Colors denote data source categories: green for fishery, orange for scientific surveys, and purple for predator-based indices. These patterns highlight spatial and temporal heterogeneity in krill dynamics across the subarea."}

@@

```{r warning=FALSE}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

dat_expanded <- dat_long %>%
  uncount(weights = Frecuencia)
```

Length compositions in Figure \@ref(fig:length)

```{r length, fig.width=10, fig.height=4, warning=FALSE, fig.cap="Annual length-frequency distributions of Antarctic krill (Euphausia superba) across different data sources and spatial strata within Subarea 48.1 from 1991 to 2020. Each panel represents a distinct stratum for either fishery-dependent (green), fishery-independent survey (orange), or predator-based (purple) observations. Density ridgelines illustrate variation in krill size structure across years. The red vertical line marks a recruit references length (3.6 cm)."}
p2 <- ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
  geom_density_ridges_gradient(scale = 1.8,
                               rel_min_height = 0.001,
                               col = "black",
                               alpha=0.4) +
  facet_wrap(~FltSvy, ncol = 10) +
  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
             color = "red", 
             size = 0.7,
             alpha = 0.5) +
  scale_fill_manual(values = c("FISHERYBS" = '#1b9e77',  
                               "FISHERYEI" = '#1b9e77',  
                               "FISHERYGS" = '#1b9e77',
                               "FISHERYJOIN" = '#1b9e77',  
                               "FISHERYSSIW" = '#1b9e77',  
                               "SURVEYBS" = '#d95f02',  
                               "SURVEYEI" = '#d95f02',  
                               "SURVEYGS" = '#d95f02',  
                               "SURVEYJOIN" = '#d95f02',  
                               "SURVEYSSIW" = '#d95f02',  
                               "PREDATOR" = '#7570b3')) +  
  labs(title = "",
       x = "cm",
       y = "",
       fill = "Category") +
  theme_bw()+
  theme(legend.position = "none")
p2
```

@@

```{r}
# Crear una columna de modelo en cada data frame
outps1$Model <- "s1.1 (Ref Model: No Env-Predator)"
outps2$Model <- "s1.2 w/ Predator data"
outps3$Model <- "s1.3 w/ Env data"
outps4$Model <- "s1.4 (Env and Predator data)"

@@

outpsall_long <- outpsall %>%
  pivot_longer(cols = c(Bio_smry, SpawnBio, Recruit_0), 
               names_to = "Variable", values_to = "Value") %>% 
  filter(Yr >= 1989, Yr <= 2020)

# Gráfico de barras agrupadas por variable y modelo
alloutput <- ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, 
             ncol=1,
             scales = "free_y") + 
  scale_fill_manual(name= "Scenario",
                    values = c("s1.1 (Ref Model: No Env-Predator)" = "#ca0020",

                               "s1.4 (Env and Predator data)" = "#404040")) +
  labs(title = "",
       x = "", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "bottom")
alloutput
```

@@

SSplotComparisons(summaryoutput1_4,
                  legendlabels = c("s1.1 (Ref Model)",
                 "s1.2",
                 "s1.3",
                 "s1.4"),
                 filenameprefix = "COM1",
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"),
                 png=TRUE,
                 plotdir=Figs)
```

@@

Antoher comparision in `R0` (Natural log of virgin recruitment level)

```{r}
r0_vals <- data.frame(
  model = c("s1.1", "s1.2", "s1.3", "s1.4"),
  value = c(base.model1.1$jitter_info[1,1],
            base.model1.2$jitter_info[2,1],
            base.model1.3$jitter_info[1,1],
            base.model1.4$jitter_info[2,1]),
  cv = c(base.model1.1$jitter_info[1,6],
         base.model1.2$jitter_info[2,6],
         base.model1.3$jitter_info[1,6],
         base.model1.4$jitter_info[2,6])
)

x_vals <- seq(10, 30, length.out = 1000)

dens_data <- r0_vals %>%
  rowwise() %>%
  mutate(sd =  cv,  
         density = list(data.frame(
           x = x_vals,
           y = dnorm(x_vals, mean = value, sd = sd),
           model = model
         ))) %>%
  select(density) %>%
  unnest(cols = c(density))

colores <- c("s1.1" = "#ca0020", 
             "s1.2" = "#f4a582", 
             "s1.3" = "#bababa", 
             "s1.4" = "#404040")

r0plot <- ggplot(dens_data, aes(x = x, y = y, color = model, fill = model)) +
  geom_area(alpha = 0.5, position = "identity") +
  geom_line(size = 1) +
  scale_color_manual(values = colores) +
  scale_fill_manual(values = colores) +
  labs(title = "",
       x = "Natural log of virgin recruitment level (R0)", y = "Density") +
  theme_bw() +
  xlim(16, 22)
r0plot
```
